\subsection{Mass as Resistance to $\chi$ Relaxation}
  \label{subsec:mass_as_resistance}

  Building on the ontological interpretation of mass as frozen structural information
  (Section~\ref{subsec:mass-as-frozen-information}), this section develops the
  quantitative formulation.
  Mass is not introduced as an intrinsic or fundamental property of matter, but
  emerges at the level of effective descriptions as a measure of how strongly a
  localized projected configuration resists admissible relaxation ordering.

  A particle-like excitation is described, within effective regimes, as a stable and
  localized projected configuration, denoted $\chi_{\mathrm{eff},s}$, characterized
  by persistent internal structure.
  Such configurations locally constrain the admissible relaxation ordering relative
  to a homogeneous effective background.
  When a geometric description applies, this constraint manifests phenomenologically
  as inertial persistence and gravitational time dilation.

  We define the effective structural energy associated with a projected solitonic
  configuration $\chi_{\mathrm{eff},s}$ as a measure of the excess resistance to
  relaxation encoded in its internal structure:
  \begin{equation}
    E[\chi_{\mathrm{eff},s}] \;\equiv\;
    \int_{\Sigma}
    \left(
      \frac{1}{\sqrt{1 - |\nabla \chi_{\mathrm{eff},s}|^2 / c^2}} - 1
    \right)
    \, d\Sigma ,
    \label{eq:chi_soliton_energy}
  \end{equation}
  where $\Sigma$ denotes a hypersurface of constant effective ordering parameter,
  and $|\nabla \chi_{\mathrm{eff},s}|$ quantifies effective structural deformation
  within the projected description.
  This expression does not represent a fundamental energy stored in the $\chi$
  substrate, but a descriptive measure of how strongly a given projected
  configuration departs from homogeneous relaxation ordering.

  The inertial mass associated with such a configuration is then defined
  operationally as
  \begin{equation}
    m \;\equiv\; \frac{E[\chi_{\mathrm{eff},s}]}{c^2}.
    \label{eq:mass_definition}
  \end{equation}

  For example, the electron's mass $m_e$ reflects its \textbf{topological stability} as a $4\pi$-periodic soliton
  (Section~\ref{subsec:4pi_soliton}), while the proton's larger mass arises from a composite 3-soliton configuration
  (Appendix~\ref{subsec:mass_as_resistance}).

  This relation is not postulated as a fundamental axiom.
  It follows directly from the role of $E[\chi_{\mathrm{eff},s}]$ as a measure of
  resistance to effective relaxation ordering.
  The universal constant $c$ appears as the maximal admissible relaxation rate and
  therefore provides the unique conversion factor between structural resistance and
  inertial response.

  Within this framework, the relation $E = mc^2$ is interpreted as a kinematic
  identity.
  Mass quantifies the amount of effective relaxation resistance locally encoded in a
  persistent projected configuration, while energy represents the same quantity
  expressed in relaxation units.

  In this sense, mass is not an independent attribute of matter.
  It is a derived and stable property of localized projected configurations,
  characterizing how strongly they resist the irreversible ordering that, in
  effective descriptions, defines physical time.

  The question of how different particle masses arise from distinct classes of
  projected configurations is addressed in
  Appendix~\ref{app:topological_solitons}, where a spectral characterization of the
  stability properties of admissible projected descriptions is proposed as the
  geometric origin of mass hierarchies.
