\subsection{Quarks as Non-Projectable Internal Modes}
  \label{subsec:quarks-non-projectable-modes}

  The ontological status of quarks provides a particularly clear illustration of the
  distinction between the pre-geometric $\chi$ substrate and its effective spacetime
  projections.

  In Cosmochrony, quarks are not interpreted as fundamental particle-like entities, nor
  as independent localized excitations of the $\chi$ substrate.
  Instead, they correspond to internal structural modes of composite solitonic
  configurations.
  These modes are required to characterize the internal organization, stability, and
  spectral structure of hadronic excitations, but do not admit an autonomous and
  coherent projection into spacetime.

  In effective quantum field descriptions, quarks appear as elementary degrees of
  freedom subject to confinement.
  Within Cosmochrony, this confinement is not imposed dynamically by an external
  interaction or force.
  Rather, it reflects a deeper structural constraint: isolated quark-like modes do not
  correspond to admissible standalone projections of $\chi_{\mathrm{eff}}$.
  Only collective configurations in which such internal modes are topologically and
  relationally closed admit a stable spacetime manifestation.

  In this sense, quarks are real at the structural level of the $\chi$ substrate, but
  incomplete at the level of physical projection.
  They are neither fictitious entities nor fundamental spacetime objects, but
  non-factorizable internal components of projected excitations.
  Their physical influence is therefore necessarily indirect, encoded in the spectral
  properties, symmetry patterns, and dynamical responses of hadrons rather than in
  localized detection events.

  This interpretation parallels the status of internal degrees of freedom in other
  collective systems: such modes are indispensable for an accurate effective
  description, yet do not correspond to independently realizable physical entities.
  Quark confinement thus appears not as a contingent feature of strong interactions,
  but as a direct and unavoidable consequence of the non-injective character of the
  projection from the $\chi$ substrate to emergent spacetime descriptions.
