\subsection{Relation to the Higgs Mechanism: Emergence from $\chi$ Dynamics}
  \label{subsec:relation-to-the-higgs-mechanism}

  In the Standard Model, the Higgs mechanism accounts for mass generation through
  spontaneous symmetry breaking of the electroweak gauge group
  $SU(2)_L \times U(1)_Y$.
  The Higgs field $\phi_H$ acquires a non-zero vacuum expectation value (VEV),
  $\langle \phi_H \rangle \simeq 246~\mathrm{GeV}$, thereby generating masses for
  fermions and gauge bosons through Yukawa and gauge couplings.

  Within the Cosmochrony framework, the Higgs field and its VEV are not regarded as
  fundamental ontological entities.
  Instead, they arise as \emph{effective low-energy descriptors} of a specific
  structural and projective regime of the underlying relational substrate $\chi$.
  Electroweak symmetry breaking and the associated mass scale are thus reinterpreted
  as emergent phenomena associated with the relaxation dynamics and topological
  organization of $\chi$, without altering the empirical content of the Standard Model.

  \subsubsection*{Structural Transition and Emergence of the Higgs VEV}
    \label{subsec:emergence-higgs-vev}

    In Cosmochrony, electroweak symmetry breaking corresponds to a structural and
    projective transition of the $\chi$ substrate between two regimes:

    \begin{itemize}
      \item \textbf{Homogeneous regime} ($\chi < \chi_c$):
      The relaxation of $\chi$ is approximately uniform and strongly constrained.
      Only highly coherent, global configurations are projectively admissible, and no
      stable, localized excitation modes are permitted.
      Effective descriptions in this regime remain massless, and the electroweak
      symmetry is unbroken.

      \item \textbf{Structured regime} ($\chi \gtrsim \chi_c$):
      As relaxation proceeds, relational constraints are sufficiently relaxed to allow
      the stabilization of localized, relaxation-resistant configurations.
      These configurations correspond to discrete, spectrally stable modes of the
      effective relaxation operator and manifest as massive excitations in spacetime-
      based descriptions.
    \end{itemize}

    This transition is not driven by an externally imposed scalar potential, but by the
    intrinsic relaxation dynamics of $\chi$.
    When the critical \emph{projective structural scale} $\chi_c$ is reached, local
    relaxation slows sufficiently to render such localized configurations
    projectively admissible.
    In effective field-theoretic language, this structural transition is described by
    the emergence of a non-zero Higgs vacuum expectation value.

  \subsubsection*{Relation Between $\chi_c$ and the Electroweak Scale}
    \label{subsec:chi_c-electroweak-scale}

    The scale $\chi_c$ should not be interpreted as a new fundamental constant.
    Rather, it represents the effective stability threshold at which localized
    configurations of $\chi$ become projectively admissible within the evolving
    configuration space.
    Its value is constrained by both cosmological relaxation properties and
    microscopic stability considerations, including:
    \begin{itemize}
      \item large-scale relaxation dynamics inferred from cosmological observations,
      \item the observed hierarchy of particle masses and stability scales.
    \end{itemize}

    At the effective level, the electroweak scale is related to $\chi_c$ through the
    inverse correlation length associated with stable $\chi$ configurations:
    \begin{equation}
      \langle \phi_H \rangle
      \;\propto\;
      \frac{\hbar_{\mathrm{eff}}\, c}{\chi_c},
    \end{equation}
    where $\hbar_{\mathrm{eff}}$ denotes the effective reprojection scale, which reduces
    to the observed Planck constant $\hbar$ in regimes where a standard quantum
    description applies.

    This relation reflects the geometric and topological conditions required for the
    stabilization of localized configurations within $\chi$, rather than the tuning of
    independent parameters.
    For $\chi_c$ of order $10^{-18}\,\mathrm{m}$, the observed electroweak scale is
    naturally recovered.

  \subsubsection*{Mass Generation as Solitonic Stabilization}
    \label{subsec:mass-generation-solitons}

    In the structured regime ($\chi \gtrsim \chi_c$), fermions and gauge bosons acquire
    mass through their association with distinct classes of stable $\chi$
    configurations.

    \begin{itemize}
      \item \textbf{Fermions:}
      Fermionic degrees of freedom correspond to topologically non-trivial,
      skyrmion-like solitonic configurations of $\chi$.
      Their effective masses scale as
      \[
        m_f \;\propto\; y_f \,\frac{\hbar_{\mathrm{eff}}}{\chi_c},
      \]
      where $y_f$ represents an effective Yukawa coupling encoding the internal
      topological and spectral properties of the configuration.
      Fermion mass hierarchies thus reflect differences in these internal invariants
      rather than independent fundamental parameters.

      \item \textbf{Gauge bosons:}
      Massive gauge bosons correspond to vortex-like or phase-structured $\chi$
      configurations.
      Their masses scale as
      \[
        m_W \;\propto\; g \,\frac{\hbar_{\mathrm{eff}}}{\chi_c},
      \]
      where $g$ is the effective $SU(2)_L$ gauge coupling.
      The weak mixing angle $\theta_W$ is interpreted as a ratio of characteristic
      topological responses associated with neutral and charged excitation sectors.
    \end{itemize}

    At the level of effective quantum field theory, these relations reproduce the
    standard Higgs-generated mass terms without modifying their phenomenology.

  \subsubsection*{Phenomenological Status and Open Questions}
    \label{subsec:phenomenological-implications}

    The emergent interpretation of the Higgs mechanism proposed here is designed to be
    phenomenologically equivalent to the Standard Model within currently tested energy
    regimes.
    No deviation from established collider results is implied at accessible energies.

    Potentially observable departures may arise only in extreme regimes, such as strong
    gravitational confinement or highly non-equilibrium relaxation, where the
    assumptions underlying an effective Higgs field description may break down.

    Open challenges include:
    \begin{itemize}
      \item deriving the detailed mapping between $\chi$ soliton spectra and the full
      Standard Model mass spectrum,
      \item understanding the origin of gauge coupling values and symmetry structure
      from the internal relational organization of $\chi$.
    \end{itemize}

  \subsubsection*{Summary}
    \label{subsec:higgs-summary}

    In Cosmochrony, the Higgs field is interpreted as an effective manifestation of a
    structured and historically admissible relaxation regime of the $\chi$ substrate.
    The electroweak scale emerges from the inverse correlation length associated with
    stable $\chi$ configurations, while particle masses arise from the topological and
    spectral stability of these configurations.

    Electroweak symmetry breaking thus corresponds not to the activation of a new
    fundamental interaction, but to the historical opening of the admissible
    configuration space allowing such localized invariants to become projectively
    stable.
    This reinterpretation preserves the empirical content of the Higgs mechanism while
    embedding it within a unified pre-geometric framework in which gravitation, quantum
    phenomena, and mass generation share a common dynamical origin.
