\subsection{Hamiltonian Derivation of the Evolution Equation}
  \label{subsec:hamiltonian-derivation}

  \subsubsection*{Local Constraint on Effective Relaxation Dynamics}

    At the fundamental level, the $\chi$ substrate does not evolve in time, admits no
    phase-space structure, and is not governed by a Hamiltonian dynamics.
    No spacetime coordinates, canonical variables, or variational principles are assumed
    in its definition.

    Nevertheless, once projected $\chi$ configurations admit a smooth and stable
    coarse-grained description, the admissible ordering of these configurations is
    subject to universal local constraints.
    These constraints may be summarized, \emph{within effective descriptions only}, by a
    compact relation formally analogous to a Hamiltonian constraint.
    This formulation is introduced solely as a descriptive parametrization of admissible
    local relaxation patterns and does not define a fundamental dynamics.

    Specifically, the effective local relaxation ordering associated with projected
    $\chi$ configurations is bounded by the invariant causal constant $c$, inherited
    from the structural bound $c_\chi$.
    When expressed in effective relational variables, this constraint takes the form
    \begin{equation}
      \left(\mathcal{D}_{\mathrm{loc}} \chi_{\mathrm{eff}}\right)^2
      + \mathcal{V}_{\chi_{\mathrm{eff}}}^2
      = c^2 ,
      \label{eq:effective_hamiltonian_constraint}
    \end{equation}
    where $\mathcal{V}_{\chi_{\mathrm{eff}}}$ denotes an effective internal variation
    functional characterizing the resistance of projected configurations to further
    relaxation.
    This quantity encodes the presence of localized structural constraints (e.g.,
    solitonic or bound configurations) and carries no interpretation as a potential
    energy in a fundamental phase space.
    No notion of spatial gradient, background geometry, or intrinsic field variation is
    assumed at this level.

    Restricting attention to the monotonic ordering branch yields the effective evolution
    relation
    \begin{equation}
      \mathcal{D}_{\mathrm{loc}} \chi_{\mathrm{eff}}
      = c \sqrt{1 - \frac{\mathcal{V}_{\chi_{\mathrm{eff}}}^2}{c^2}} ,
      \label{eq:chi_dynamics}
    \end{equation}
    which expresses the universal slowdown of effective relaxation induced by localized
    structural constraints.
    The term $\mathcal{V}_{\chi_{\text{eff}}}$ encodes \textbf{structural resistance} to relaxation from localized
    configurations (e.g., solitons), not a potential energy in a fundamental phase space.
    This relation constrains admissible projected descriptions but does not represent an
    equation of motion for the fundamental $\chi$ substrate.

  \subsubsection*{Emergent Gravitational Description}

    Projected $\chi$ configurations exhibiting strong resistance to relaxation correspond,
    in effective descriptions, to localized regions of enhanced structural complexity.
    Such regions locally reduce the admissible effective relaxation rate.

    When a spacetime interpretation becomes applicable, this reduction is conventionally
    described as gravitational time dilation.
    No independent gravitational field or interaction is postulated.
    The phenomenon arises directly from the constrained ordering of projected $\chi$
    configurations imposed by the saturation condition
    in Eq.~\eqref{eq:effective_hamiltonian_constraint}.

    In weak-structure regimes, where internal variation rates are small compared to $c$,
    the effective description admits a simplified relation governing the spatial
    distribution of relaxation slowdown:
    \begin{equation}
      \nabla \cdot \left(
                     \frac{\nabla \chi_{\mathrm{eff}}}
                     {\sqrt{1 - |\nabla \chi_{\mathrm{eff}}|^2 / c^2}}
      \right)
      \simeq \frac{4 \pi G_{\mathrm{eff}}}{c^2} \rho ,
      \label{eq:effective_poisson_relation}
    \end{equation}
    where $\rho$ denotes the effective density of localized relaxation-resistant
    configurations and $G_{\mathrm{eff}}$ is an emergent coupling parameter characterizing
    the collective response of the relaxation ordering to such structures.
    Both quantities are defined strictly at the effective level.

    In the Newtonian limit, this relation reduces to an effective Poisson equation for a
    potential $\Phi$, defined operationally by
    \begin{equation}
      \Phi \equiv c^2 \ln \!\left(
                              \frac{\mathcal{D}_{\mathrm{loc}} \chi_{\mathrm{eff}}}{c}
      \right) .
      \label{eq:effective_gravitational_potential}
    \end{equation}
    This potential is not introduced as a fundamental field, but as a compact summary of
    how localized variations in effective relaxation ordering modulate physical clocks
    and rulers in regimes where classical gravitational phenomenology applies.

  \subsubsection*{Interpretational Status}

    The relations presented in this subsection do not define a fundamental Hamiltonian,
    action, or variational principle.
    They provide an effective local parametrization of the structural constraints
    governing admissible projected descriptions once a geometric interpretation becomes
    meaningful.

    The predictive content of Cosmochrony resides entirely in the structural properties
    of the pre-geometric $\chi$ substrate.
    Hamiltonian, geometric, and gravitational formalisms arise only as emergent
    descriptive tools, valid in restricted regimes and carrying no independent
    ontological status.
