\subsection{Microscopic Origin of the Coupling Tensor and the Poisson Equation}
  \label{subsec:microscopic-origin-of-the-coupling-tensor-and-the-poisson-equation}

  For internal consistency, the effective coupling governing the relaxation ordering of
  projected $\chi$ configurations cannot be treated as a fixed universal constant.
  Instead, it must depend on the internal structural state of the projected description,
  reflecting how localized configurations resist or facilitate further relaxation.
  In Cosmochrony, this dependence is captured through a constitutive relation linking
  effective coupling strengths to variations of the effective scalar descriptor
  $\chi_{\mathrm{eff}}$, without invoking any underlying spatial substrate at the
  fundamental level.

  A convenient phenomenological parametrization of this dependence is given by
  \begin{equation}
    K_{\mathrm{eff}}
    = K_0 \exp\!\left(
                  -\frac{(\Delta \chi_{\mathrm{eff}})^2}{\chi_c^2}
    \right),
    \label{eq:effective_coupling_tensor}
  \end{equation}
  where $\Delta \chi_{\mathrm{eff}}$ denotes a measure of effective internal variation
  between correlated projected configurations,
  $K_0$ characterizes the maximal relaxation conductivity in a homogeneous effective
  background, and $\chi_c$ sets the characteristic scale beyond which structural
  inhomogeneities significantly reduce the effectiveness of relaxation.
  This parametrization is purely phenomenological and does not represent a fundamental
  law at the level of the $\chi$ substrate.

  Projected configurations exhibiting strong effective internal variation—such as
  stable solitonic structures—therefore reduce the effective coupling and locally slow
  the admissible relaxation ordering.
  This reduction does not correspond to an additional interaction, but reflects the
  intrinsic resistance of structured projected configurations to further relaxation.
  The resulting slowdown constitutes the microscopic origin of the emergent
  gravitational phenomenology discussed in the previous sections.

  In regimes where a spacetime description becomes applicable, the local effective
  relaxation rate $\mathcal{D}_{\mathrm{loc}} \chi_{\mathrm{eff}}$ differs from its
  asymptotic value $\mathcal{D}_0$ far from localized structures.
  An effective gravitational potential $\Phi$ may then be introduced as a descriptive
  parameter through the relation
  \begin{equation}
    \frac{\mathcal{D}_{\mathrm{loc}} \chi_{\mathrm{eff}}}{\mathcal{D}_0}
    \simeq 1 + \frac{\Phi}{c^2},
    \label{eq:relaxation_potential_relation}
  \end{equation}
  which summarizes the relative slowdown of effective relaxation ordering in a form
  familiar from classical gravitational phenomenology.

  In the weak-structure regime, where effective internal variations remain small
  compared to $\chi_c$, the spatial distribution of $\Phi$ admits a simplified elliptic
  description.
  At this coarse-grained level, the effective dynamics reduce to a Poisson-type
  relation,
  \begin{equation}
    \nabla^2 \Phi \simeq 4\pi G_{\mathrm{eff}} \rho,
    \label{eq:effective_poisson_equation}
  \end{equation}
  where $\rho$ denotes the effective density of localized, relaxation-resistant
  projected configurations and $G_{\mathrm{eff}}$ is an emergent coupling parameter
  encoding the collective response of the effective relaxation ordering to such
  structures.

  This Poisson equation is not fundamental.
  It represents the weak-field, macroscopic limit of the constrained ordering of
  projected $\chi$ configurations, expressed in a form adapted to effective geometric
  description.
  Gravitation therefore appears not as an independent interaction, but as a descriptive
  manifestation of reduced relaxation conductivity induced by structured projected
  configurations.

  A fully relational formulation, consistent with but not required for the effective
  description adopted here, is provided in
  Appendix~\ref{app:relational_formulation}.

  \paragraph{Status of effective equations.}
    The effective equations introduced in this section—such as the Poisson-like relation
    for the gravitational potential or the wave equations describing solitonic
    structures—are valid only within regimes where projected $\chi$ configurations admit
    a smooth, weak-gradient geometric interpretation.
    They arise from coarse-graining admissible projected descriptions and do not
    represent fundamental postulates of the theory.

    In particular, these equations should not be confused with any intrinsic relaxation
    law governing the $\chi$ substrate at the pre-geometric level.
    Their domain of applicability is limited to configurations close to local equilibrium,
    where an effective spacetime description becomes meaningful.
