\subsection{Causality and Locality}
  \label{subsec:causality-and-locality}

  Equation~\eqref{eq:chi_dynamics} does not define a fundamental dynamics of the
  $\chi$ substrate.
  Rather, it constrains admissible projected descriptions in regimes where a
  geometric interpretation becomes applicable.
  Within such effective descriptions, the relaxation ordering encoded by
  $\chi_{\mathrm{eff}}$ exhibits locality and causality in the operational sense
  relevant to physical observables.

  Effective locality follows from the fact that variations of
  $\chi_{\mathrm{eff}}$ at a given effective spacetime event depend only on
  correlated neighboring projected configurations, as defined within the emergent
  geometric description.
  No instantaneous coupling, action at a distance, or nonlocal influence is
  introduced at the level of effective physical observables.
  All admissible projected descriptions therefore respect locality in the sense
  appropriate to an emergent spacetime framework.

  Causality is enforced through the existence of a universal bound on the effective
  local relaxation ordering,
  \[
    \left| \mathcal{D}_{\mathrm{loc}} \chi_{\mathrm{eff}} \right| \le c ,
  \]
  which constrains the maximal rate at which correlations between effective
  configurations may be established.
  This bound functions as an effective causal constraint without presupposing a
  fundamental spacetime structure, lightcones, or signal propagation at the level
  of the $\chi$ substrate.

  Within effective physical descriptions, no superluminal propagation of signals
  or causal influence occurs.
  Apparent superluminal recession velocities observed in cosmological contexts
  arise solely from the cumulative integration of locally constrained relaxation
  ordering over extended regions.
  They therefore remain fully consistent with effective locality and causality as
  defined within the Cosmochrony framework.

  \medskip
  \noindent\emph{Conceptual remark.}—
  The effective causal bound $c$ introduced here should not be interpreted as an
  independent physical postulate.
  As established in Section~\ref{subsec:role-of-cchi}, it corresponds to the
  projected manifestation of the invariant structural bound $c_\chi$ defined at
  the level of the pre-temporal $\chi$ substrate.
  In regimes admitting a locally injective projection and a stable geometric
  interpretation, $c_\chi$ acquires an operational meaning as a maximal admissible
  local ordering rate, which appears in spacetime descriptions as the effective
  causal constraint $c$.

  Effective causality in Cosmochrony thus arises as a derived property of bounded
  projective realizations, rather than as a fundamental principle imposed on
  spacetime itself.
