\subsection{Unified Origin of Geometric and Field Effects}
  \label{subsec:unified-origin}

  The relation between the $\chi$ substrate and the effective spacetime metric
  $g_{\mu\nu}$ is strictly hierarchical.
  It reflects the transition from a fundamental pre-geometric relational structure
  to smooth geometric descriptions applicable only at macroscopic scales.

  \begin{enumerate}
    \item \textbf{Primacy of the $\chi$ Substrate.}
    At the fundamental level, physical reality is described solely in terms of the
    $\chi$ substrate and its intrinsic relational structure.
    The $\chi$ substrate is not defined on spacetime, does not possess numerical
    values, and is not governed by a dynamical field equation.
    Ordering, relaxation, and causal notions arise only within admissible projected
    descriptions and have no meaning at the pre-geometric level.

    \item \textbf{Emergent Geometry.}
    In regimes where projected $\chi$ configurations admit a stable and slowly varying
    description, geometric notions become meaningful.
    The spacetime metric $g_{\mu\nu}$ arises as an effective descriptor summarizing
    correlations and relaxation ordering among projected $\chi$ configurations.
    It provides a coarse-grained geometric language suitable for macroscopic
    descriptions, without acquiring independent ontological or dynamical status.

    \item \textbf{Unified Interpretation of Fields and Gravitation.}
    Within the effective geometric description, localized relaxation-resistant
    projected configurations—describable as solitonic structures—are identified with
    matter degrees of freedom.
    Gravitational phenomena correspond to local modulations of effective relaxation
    ordering induced by such structures.
    The metric does not act as an independent dynamical agent; it encodes the
    collective geometric response associated with constrained projected descriptions.
  \end{enumerate}

  In this framework, no independent gravitational interaction or fundamental field is
  postulated.
  Matter, geometry, and gravitation emerge as complementary and inseparable aspects of
  the same constrained ordering of projected $\chi$ configurations.
  This unified origin ensures conceptual continuity across scales, from localized
  particle-like structures to macroscopic spacetime geometry.
