\subsection{The Geometric Effective Description of $\chi$ Dynamics}
  \label{subsec:geometric-action}

  \subsubsection*{Effective Observables from $\chi$ Correlations}

    In Cosmochrony, quantities conventionally described in geometric terms—such as time
    intervals, spatial separation, and causal ordering—are not taken as primitive.
    They arise as effective descriptive summaries of relational patterns within the
    $\chi$ substrate, accessed only through projected, coarse-grained representations.
    No background spacetime, coordinate system, or discrete substrate is assumed at any
    stage of the fundamental description.

    The configurations $\sigma$ represent internal states of the $\chi$ relational substrate
    and are defined without reference to external spacetime coordinates or background
    geometry.
    They label relational degrees of freedom of the substrate, specifying patterns of
    internal organization rather than positions in a pre-existing space.

    Correlations between configurations $\sigma$ encode the emergent geometric and causal
    structure at the effective level, once a stable spacetime description becomes
    applicable.
    The measure $d\mu(\sigma)$ denotes an invariant integration over this space of
    configurations, defined intrinsically from the correlation structure associated with
    $\chi$.
    It ensures that physical observables are independent of the particular parametrization
    chosen to label configurations, and carries no interpretation as a volume element in an
    underlying spacetime.

    \paragraph{Effective scalar descriptor.}
      In regimes where projected $\chi$ configurations admit a stable geometric
      interpretation, it is convenient to introduce an \emph{effective scalar descriptor},
      denoted $\chi_{\mathrm{eff}}$.
      This quantity is a coarse-grained, projected representation of relational features of
      the $\chi$ substrate, defined only within the emergent spacetime description.
      It does not correspond to the fundamental $\chi$ substrate itself, which remains
      non-indexable and devoid of intrinsic values.
      Although $\chi_{\mathrm{eff}}$ is represented as a scalar function in effective
      descriptions, it should not be interpreted as a fundamental field propagating in
      spacetime.

      Operational time intervals are defined from the accumulated ordering of projected
      $\chi$ configurations along paths in configuration space.
      This ordering is quantified using the effective descriptor $\chi_{\mathrm{eff}}$, not
      the fundamental $\chi$ substrate:
      \begin{equation}
        \tau_{AB} \;\propto\;
        \int_{\gamma_{AB}} \mathcal{D}_{\lambda} \chi_{\mathrm{eff}} \, d\lambda,
      \end{equation}
      where $\mathcal{D}_{\lambda} \chi_{\mathrm{eff}}$ is an effective relaxation functional
      characterizing the ordering of projected $\chi$ configurations along the path
      $\gamma_{AB}$.
      The parameter $\lambda$ is an ordering parameter, not a fundamental time coordinate.

      Geometric observables are constructed from correlations between effective scalar
      descriptors $\chi_{\mathrm{eff}}$, associated with projected $\chi$ configurations.
      These correlations encode relational information about causal connectivity and
      separation once a geometric regime is established.

      Operational spatial separation is quantified by the decay of correlations between
      effective scalar descriptors $\chi_{\mathrm{eff}}$ associated with projected $\chi$
      configurations:
      \begin{equation}
        d(x,y) \;\propto\;
        - \log\left(
                \frac{\langle \chi_{\mathrm{eff}}(x)\,\chi_{\mathrm{eff}}(y)\rangle}
                {\langle \chi_{\mathrm{eff}}^{\,2}\rangle}
        \right),
      \end{equation}
      where $\langle \chi_{\mathrm{eff}}(x)\chi_{\mathrm{eff}}(y)\rangle$ measures the
      correlation between effective descriptors associated with configurations labeled by
      $x$ and $y$.
      Here $x$ and $y$ label effective spacetime events in the emergent description.
      The correlation function involves $\chi_{\mathrm{eff}}$ only and does not imply any
      localization or value assignment at the level of the fundamental $\chi$ substrate.

      These definitions are purely relational and make no reference to a pre-existing metric
      or discrete structure.
      They are applicable whenever projected $\chi$ configurations exhibit stable
      correlation patterns.

      Throughout this section, all indexed, correlated, or integrated quantities refer
      exclusively to the effective descriptor $\chi_{\mathrm{eff}}$, not to the fundamental
      $\chi$ substrate itself.

  \subsubsection*{Effective Metric as a Descriptive Tool}

    In regimes where projected $\chi$ configurations exhibit smooth and stable correlation
    patterns, the relational observables defined above may be compactly summarized by an
    operational tensor $g_{\mu\nu}[\chi_{\mathrm{eff}}]$.
    The notation $g_{\mu\nu}[\chi_{\mathrm{eff}}]$ emphasizes that the metric summarizes
    correlations of effective descriptors, not properties of the fundamental $\chi$
    substrate.
    This object is not introduced as a fundamental geometric structure, nor as an
    independent degree of freedom.
    Rather, it provides a convenient parametrization of how variations in
    $\chi_{\mathrm{eff}}$ modulate causal connectivity and effective relational intervals,
    and is meaningful only insofar as such a coarse-grained description remains valid.

    The metric is introduced as a descriptive summary of $\chi_{\mathrm{eff}}$
    correlations, not postulated as an independent degree of freedom.
    For example:
    \begin{itemize}
      \item The conformal (lightcone) structure is constrained by the maximal effective
      relaxation speed $c$ associated with $\chi_{\mathrm{eff}}$.
      \item Proper time between effective events is proportional to the accumulated
      $\chi_{\mathrm{eff}}$ relaxation along paths connecting the corresponding
      configurations.
      \item Spatial distance reflects the decay rate of $\chi_{\mathrm{eff}}$
      correlations.
    \end{itemize}

    No discrete-to-continuum limit is invoked.
    The theory is continuous at all scales; apparent granularity (e.g., Planck-scale
    phenomena) is attributed to non-linear $\chi$ dynamics rather than to an underlying
    discretization.

    No background $\eta_{\mu\nu}$ is assumed.
    Minkowski space appears only as a convenient approximation in suitable limits (e.g.,
    weak-gradient regimes), without ontological status in the fundamental description.

  \subsubsection*{Consistency with General Relativity}

    The effective metric $g_{\mu\nu}[\chi_{\mathrm{eff}}]$, constructed as a summary of
    $\chi_{\mathrm{eff}}$ correlations, reproduces the phenomenology of general relativity
    in the following sense:
    \begin{itemize}
      \item \textbf{Weak-field limit:} when $\chi_{\mathrm{eff}}$ gradients are small, the
      effective metric approaches a form compatible with Einstein-like dynamics for a
      fluid-like stress-energy description associated with $\chi$ excitations.
      \item \textbf{Strong-field regimes:} near localized $\chi$ excitations (e.g.,
      solitons), the metric encodes time dilation and spatial curvature as emergent
      effects of slowed $\chi_{\mathrm{eff}}$ relaxation, without requiring a
      fundamental gravitational field.
      \item \textbf{Cosmological expansion:} homogeneous relaxation of $\chi$ yields an
      effective Hubble-like expansion law for the emergent scale factor.
    \end{itemize}

    Crucially, this is not a bootstrap procedure.
    The metric is not iteratively reconstructed from $\chi$; it is a post-hoc descriptive
    tool summarizing geometric regularities of projected $\chi$ configurations.
    The theory's predictive content resides in the dynamics of $\chi$, not in the metric
    itself.

  \subsubsection*{Ontological Status of the Metric}

    To avoid confusion, we emphasize:
    \begin{itemize}
      \item $\chi$ is the only fundamental entity.
      Spacetime, metric structure, and matter are emergent descriptions of projected
      $\chi$ configurations.
      \item No ``double ontology'' is assumed: there is no underlying discrete graph or
      lattice.
      Geometric language is introduced only as an effective descriptive tool.
      \item The metric $g_{\mu\nu}[\chi_{\mathrm{eff}}]$ is an effective construct, analogous
      to how temperature emerges in thermodynamics.
      It is useful for coarse-grained description but plays no role in the fundamental
      dynamics.
    \end{itemize}

    \paragraph{Operational origin of the effective metric.}
      The metric tensor $g_{\mu\nu}$ is not a fundamental structure but a derived descriptor,
      summarizing the relational distance induced by $\chi$ correlations. Its explicit
      construction from operational distances is given in Appendix~E, where geometric
      quantities are shown to arise only in projectable regimes admitting a smooth
      continuum approximation.

    \paragraph{Operational interpretation of the line element.}
      In Cosmochrony, the line element $ds^2$ is \emph{not} a primitive geometric quantity.
      It is an effective descriptor that becomes meaningful only in projectable regimes where
      the relational distance structure admits a smooth local approximation.

      Concretely, the fundamental object is the operational distance $d_{ij}$ defined on the
      relational network of $\chi$ (Appendix~E). Once a low-dimensional embedding exists,
      the squared distance between nearby embedded points can be locally approximated by
      \[
        d(i,j)^2 \;\approx\; g_{\mu\nu}(x)\,\Delta x^\mu \Delta x^\nu,
      \]
      thereby introducing $g_{\mu\nu}$ as a \emph{derived quadratic form} summarizing
      relational connectivity. The continuum line element $ds^2$ should therefore be
      understood as a shorthand for this local approximation, not as a fundamental structure.

  \subsubsection*{Summary: A Fully Continuous Framework}

    \begin{itemize}
      \item \textbf{Fundamental level:} only the continuous $\chi$ relational substrate
      exists, characterized by an intrinsic ordering structure.
      \item \textbf{Effective level:} geometric observables (time, distance, metric) emerge
      from correlations between effective descriptors $\chi_{\mathrm{eff}}$
      associated with projected $\chi$ configurations.
      \item \textbf{No bootstrap:} the metric is never iteratively constructed or assumed; it
      is a derived description summarizing the relational structure of projected
      $\chi$ configurations.
      \item \textbf{No discretization:} apparent ``Planck-scale'' effects arise from
      non-linear $\chi$ dynamics, not from an underlying discrete substrate.
    \end{itemize}
