\clearpage
\section{Definition and Fundamental Properties of the \texorpdfstring{$\chi$}{χ} Substrate}
  \label{sec:definition-and-fundamental-properties-of-the-chi-field}

  Having outlined the conceptual motivations underlying Cosmochrony, we now introduce
  the single fundamental entity at the core of the framework.
  This section provides the \emph{canonical ontological definition} of the pre-geometric
  relational substrate $\chi$, which constitutes the primitive basis from which effective
  notions of spacetime, dynamics, and physical observables emerge.

  The aim of this section is to identify the \emph{minimal structural properties}
  required of $\chi$ for such effective descriptions to arise in appropriate regimes.
  No spacetime manifold, metric structure, or background geometry is assumed at this
  level.
  Instead, geometric and dynamical notions are recovered only through non-injective,
  coarse-grained projections of $\chi$ configurations once suitable stability conditions
  are met.
  This minimality principle, referred to throughout the paper as a regime of
  \emph{ontological poverty}, is adopted here as a foundational constraint rather than
  as a derived result.

  Throughout this section, the use of variational principles, Lagrangian formulations,
  or metric-based expressions does not imply that spacetime or a four-dimensional
  manifold is fundamental.
  Such formalisms are introduced strictly as effective descriptive tools, valid only in
  projectable regimes where configurations of $\chi$ admit a quasi-stable geometric and
  causal interpretation.
  They should therefore be understood as post-hoc representations of underlying
  pre-geometric dynamics, not as primary postulates of the theory.

  We begin by introducing the unified conceptual definition of the $\chi$ substrate and
  its physical interpretation, emphasizing its ontological distinction from conventional
  fields.
  The subsequent subsections progressively introduce more structured effective
  descriptions—including scalar, Lagrangian, and metric formulations—which become
  applicable only once projected $\chi$ configurations support a coherent spacetime
  regime.

  \input{03-chi-field/01-definition-and-fundamental-properties-of-the-chi-field}
  \input{03-chi-field/02-the-geometric-effective-action-and-lagrangians-of-cosmochrony-Lcc}
  \input{03-chi-field/03-physical-interpretation}
  \input{03-chi-field/04-relational_projection}
  \input{03-chi-field/05-monotonicity-and-arrow-of-time}
  \input{03-chi-field/06-local-relaxation-speed}
  \input{03-chi-field/07-relation-to-conventional-fields}
  \input{03-chi-field/08-initial-conditions-and-global-structure}
