\subsection{Absence of Dark Energy Signatures}
  \label{subsec:absence-of-dark-energy-signatures}

  Because cosmic acceleration emerges in Cosmochrony as a geometric consequence of
  the global relaxation of the $\chi$ field, no independent dark energy component is
  introduced.
  Accordingly, the framework predicts the absence of signatures associated with
  dynamical dark energy, such as an evolving equation of state, clustering behavior,
  or additional propagating degrees of freedom beyond those already present in the
  effective geometric description.

  Within this perspective, observations consistent with a purely geometric and
  kinematic origin of late-time acceleration would favor Cosmochrony over models
  requiring additional energy components or fine-tuned scalar fields.

  \paragraph{Discriminating observational signatures.}
    The absence of dark energy dynamics cannot be established through any single
    observable.
    Instead, Cosmochrony predicts a correlated pattern of large-scale cosmological
    features reflecting the lack of an inflationary phase and the pre-geometric origin
    of early-time correlations.

    These features include suppressed power at low CMB multipoles, specific angular
    correlations in temperature and polarization, and the absence of an inflationary
    tensor imprint at large angular scales.
    It is the combined presence of these signatures—rather than any individual
    parameter—that provides a potential observational discriminator with respect to
    standard inflationary and dark-energy-driven cosmological models.
