\subsection{The Big Bang as a Maximal Constraint Regime of the $\chi$ Substrate}
  \label{subsec:big-bang-maximal-constraint}

  Within the Cosmochrony framework, the Big Bang is not interpreted as a spacetime singularity, nor as a physical event
  occurring at a definite moment.
  It corresponds instead to a limiting regime in which the relational structure of the $\chi$ substrate is maximally
  constrained.

  In this regime, the density of structural and topological constraints within $\chi$
  is such that no stable geometric projection is admissible~\cite{Einstein1915,Weinberg1972}.
  Concepts such as spatial distance, temporal duration, curvature, or causal ordering are therefore undefined.
  The apparent singular behavior encountered in standard cosmological models reflects the breakdown of effective
  spacetime descriptions when extrapolated beyond their domain of validity.

  Cosmic evolution is interpreted as the progressive relaxation of these maximal constraints.
  As the relational structure of $\chi$ becomes less constrained, increasingly stable projected descriptions become
  admissible, allowing effective notions of space, time, and geometry to emerge~\cite{Penrose1989Weyl}.
  The Big Bang thus marks not the beginning of spacetime, but the boundary beyond which spacetime ceases to be a
  meaningful descriptive framework.

  Within this perspective, the arrow of time does not originate from special initial conditions or entropy
  assumptions~\cite{Prigogine1997,Rovelli1991}.
  It arises intrinsically from the monotonic relaxation ordering of $\chi$away from the maximally constrained regime.
  Temporal ordering is therefore a consequence of the structural
  evolution of the substrate itself, rather than a feature imposed at the level of effective cosmological descriptions.
