\subsection{Coupling with Matter: Effective Source Term $S[\chi,\rho]$}
  \label{subsec:coupling_matter_chi}

  In regimes where the $\chi$ field admits a smooth geometric interpretation, its
  dynamics may be expressed using effective differential operators familiar from
  field theory.
  Within this \emph{emergent} description, the influence of localized excitations
  (matter or energy density $\rho$) on the relaxation of $\chi$ can be summarized
  by an effective source term $S[\chi,\rho]$:
  \begin{equation}
    \square_{\text{eff}} \chi = S[\chi,\rho].
  \end{equation}
  This equation should not be interpreted as fundamental.
  Both the operator $\square_{\text{eff}}$ and the source term $S$ arise only after
  a spacetime description has emerged from the underlying relaxation dynamics of
  $\chi$.

  \subsubsection{Physical Meaning of $S[\chi,\rho]$}

    The term $S[\chi,\rho]$ does not represent an external force acting on $\chi$.
    Rather, it encodes the \emph{effective resistance} of localized excitations to
    the global relaxation of the field.
    Regions containing matter correspond to structured configurations of $\chi$
    (solitons) that locally reduce the relaxation rate, inducing spatial gradients
    and differential proper-time flow.

    Within this interpretation, $S[\chi,\rho]$ provides a compact phenomenological
    description of several emergent effects:
    \begin{itemize}
      \item gravitational time dilation as a consequence of slowed $\chi$ relaxation,
      \item inertial mass as persistent resistance to relaxation,
      \item effective spacetime curvature as a coarse-grained description of spatial
      variations in $\chi$.
    \end{itemize}

  \subsubsection{Effective Form and Weak-Field Limit}

    In weak-field regimes, where matter-induced gradients are small, the effective
    source term may be approximated as linear in the excitation density:
    \begin{equation}
      S[\chi,\rho] \simeq -\alpha \rho ,
    \end{equation}
    with $\alpha$ an effective coupling constant.
    Matching with the Newtonian limit identifies $\alpha \sim G/c^2$, where $G$
    emerges as the macroscopic coupling between matter density and relaxation
    slowdown.

    This linear approximation is sufficient to recover the Poisson equation for the
    effective gravitational potential and the Schwarzschild solution at leading
    order.

  \subsubsection{Strong-Field Regimes and Nonlinear Corrections}

    In regimes of high excitation density, such as near compact objects or in the
    early universe, nonlinear corrections to $S[\chi,\rho]$ are expected.
    These corrections reflect saturation effects in the relaxation dynamics and
    prevent unphysical halting of the field evolution:
    \begin{equation}
      S[\chi,\rho] = -\alpha \rho \, F\!\left(\frac{\rho}{\rho_c}, \chi\right),
    \end{equation}
    where $F$ is a bounded function and $\rho_c$ is a characteristic density scale.
    Such nonlinearities encode departures from classical gravity without modifying
    the underlying ontological structure.

\subsection{Strong-Field Constitutive Coupling Near a Schwarzschild Black Hole}
  \label{app:keff_schwarzschild}

  \paragraph{Purpose and status.}
    Section~\ref{subsec:microscopic-origin-coupling} introduced an effective constitutive
    relation for the relaxation conductivity,
    \begin{equation}
      K_{\mathrm{eff}} = K_0 \exp\!\left(-\frac{(\Delta\chi)^2}{\chi_c^2}\right),
      \label{eq:keff_constitutive}
    \end{equation}
    as a coarse-grained way of encoding how strong internal structure of $\chi$ reduces
    the effectiveness of relaxation.
    In weak-field regimes this leads to a Poisson-like description and to the recovery of
    Schwarzschild phenomenology at leading order (Section~\ref{subsec:recovery_schwarzschild}).
    The goal of this appendix is to make explicit a consistent \emph{strong-field}
    profile $K_{\mathrm{eff}}(r)$ in the spherically symmetric case, suitable for describing
    the approach to an effective horizon (Section~\ref{subsec:strong_gravity_black_holes}).

  \paragraph{Operational time-dilation factor.}
    In the emergent geometric regime, gravitational time dilation is encoded by a local
    slowdown of the relaxation rate of $\chi$ relative to its asymptotic value far from the
    source. We define the dimensionless lapse-like factor
    \begin{equation}
      N(r) \;\equiv\; \frac{D_{\mathrm{loc}}\chi(r)}{D_0\chi},
      \qquad 0 < N(r) \le 1,
      \label{eq:lapse_def}
    \end{equation}
    where $D_0\chi$ denotes the asymptotic relaxation rate in a homogeneous background.
    In the weak-field limit, one may write $N \simeq 1+\Phi/c^2$ for an effective Newtonian
    potential $\Phi$ (Section~\ref{subsec:microscopic-origin-coupling}).

  \paragraph{Matching to Schwarzschild form.}
    Section~\ref{subsec:recovery_schwarzschild} argues that, in regimes where a stable
    geometric description applies, the external field of an isolated compact source may be
    summarized by a Schwarzschild-like line element. We adopt the standard form
    \begin{equation}
      ds^2 = -f(r)c^2 dt^2 + f(r)^{-1} dr^2 + r^2 d\Omega^2,
      \qquad f(r)=1-\frac{r_s}{r},
      \label{eq:schwarzschild_standard}
    \end{equation}
    with $r_s = 2GM/c^2$ defined operationally by the asymptotic weak-field matching.
    Consistency with the interpretation of $N(r)$ as the local time-dilation factor implies
    \begin{equation}
      N(r)^2 = f(r) = 1-\frac{r_s}{r}.
      \label{eq:lapse_schwarzschild}
    \end{equation}
    Thus $N(r)\to 0$ as $r\to r_s^+$, capturing the asymptotic freeze-out of local relaxation
    identified with an effective horizon.

  \paragraph{From lapse to strong-field conductivity.}
    To connect $N(r)$ to $K_{\mathrm{eff}}(r)$ we need a strong-field identification between the
    relaxation slowdown and the reduction of conductivity.
    A minimal and self-consistent choice is to assume that the \emph{fractional} slowdown
    is directly controlled by the \emph{fractional} conductivity,
    \begin{equation}
      \frac{K_{\mathrm{eff}}(r)}{K_0} \;\equiv\; N(r)^2,
      \label{eq:keff_lapse_ansatz}
    \end{equation}
    which (i) reproduces the weak-field expansion to leading order,
    (ii) ensures $K_{\mathrm{eff}}\to 0$ at the horizon (no relaxation transport across an
    asymptotically frozen region), and (iii) remains bounded and monotone.

    Combining \eqref{eq:keff_lapse_ansatz} with \eqref{eq:lapse_schwarzschild} yields an explicit
    strong-field profile:
    \begin{equation}
      K_{\mathrm{eff}}(r)
      \;=\;
      K_0\!\left(1-\frac{r_s}{r}\right),
      \qquad r>r_s.
      \label{eq:keff_schwarzschild_profile}
    \end{equation}
    This expression should be read as an \emph{effective constitutive law} in the emergent
    geometric regime; it does not assert that $K_{\mathrm{eff}}$ is fundamental.

  \paragraph{Implied structural variation $\Delta\chi(r)$.}
    Using the constitutive relation \eqref{eq:keff_constitutive} together with
    \eqref{eq:keff_lapse_ansatz}, one obtains the corresponding strong-field variation measure:
    \begin{equation}
      \frac{(\Delta\chi(r))^2}{\chi_c^2}
      \;=\;
      -\ln\!\left(\frac{K_{\mathrm{eff}}(r)}{K_0}\right)
      \;=\;
      -\ln\!\left(1-\frac{r_s}{r}\right),
      \label{eq:deltachi_schwarzschild}
    \end{equation}
    so that $\Delta\chi(r)$ diverges logarithmically as $r\to r_s^+$,
    \begin{equation}
      \Delta\chi(r) \sim \chi_c \sqrt{-\ln\!\left(1-\frac{r_s}{r}\right)}.
      \label{eq:deltachi_horizon_asymptotic}
    \end{equation}
    This divergence should not be interpreted as a spacetime singularity.
    It reflects that the coarse-grained structural measure $\Delta\chi$ ceases to remain small and that the
    effective geometric parametrization is pushed to its limit of validity near the horizon.

  \paragraph{Interpretation: horizons as vanishing relaxation conductivity.}
    Equations \eqref{eq:keff_schwarzschild_profile}--\eqref{eq:deltachi_schwarzschild} provide a
    compact strong-field completion of the weak-field Poisson description:
    a Schwarzschild horizon corresponds to a \emph{conductivity zero} of the relaxation flow,
    $K_{\mathrm{eff}}\to 0$, rather than to a fundamental geometric singularity.
    In this sense, black holes are regions where the collective structural constraints in $\chi$
    asymptotically inhibit relaxation, producing the temporal and gravitational shadows discussed
    in Section~\ref{subsec:strong-gravity-and-black-holes}.

\subsubsection{Scope and Open Questions}

  The effective description provided by $S[\chi,\rho]$ raises several open issues:
  \begin{itemize}
    \item the microscopic origin of the coupling constant $\alpha$,
    \item the role of $S[\chi,\rho]$ in quantum regimes where $\rho$ is replaced by
    excitation amplitudes,
    \item possible observational signatures arising from nonlinear relaxation
    effects.
  \end{itemize}

  These questions are deferred to future work.
  The present formulation is intended as a bridge between the fundamental
  relaxation dynamics of $\chi$ and the effective gravitational and cosmological
  phenomena observed at macroscopic scales.

\subsubsection{Conclusion}

  The effective source term $S[\chi,\rho]$ provides a consistent and economical way
  to encode the influence of localized excitations on the relaxation of the $\chi$
  field.
  While not fundamental, it allows Cosmochrony to recover known gravitational
  phenomena and to articulate clear predictions in regimes where deviations from
  standard theories may arise.
