\subsection{Measurement, Decoherence, and Apparent Collapse}
  \label{subsec:measurement-and-decoherence}

  Within the Cosmochrony framework, quantum measurement does not involve a fundamental
  wavefunction collapse.
  At the fundamental level, no discontinuous update of the underlying relational
  substrate occurs.
  What is conventionally described as collapse arises entirely within the domain of
  effective projected descriptions.

  Measurement corresponds to the transition from a non-factorizable admissible
  projected description to a set of effectively factorized local projections.
  This transition is induced by interactions with an environment that progressively
  eliminate the accessibility of global relational coherence.
  As a result, alternative relational components cease to admit a joint effective
  description.

  Decoherence therefore does not represent a postulated measurement axiom, but a
  dynamical restriction on admissible projected descriptions~\cite{Zurek2003}.
  It suppresses interference between incompatible descriptive branches by rendering
  their relative phase information inaccessible within spacetime representations.
  The underlying relational structure remains globally well defined, even though it
  can no longer be represented coherently at the effective level.

  Importantly, decoherence does not destroy information.
  It limits the projectability of relational correlations into spacetime descriptions.
  In this sense, decoherence may be understood as a local and partial loss of
  projectability: certain relational distinctions persist structurally but cease to be
  jointly representable.

  More extreme regimes, such as those associated with strong gravitational confinement,
  represent a limiting case of this mechanism.
  In such regimes, not only coherence but spacetime representability itself breaks
  down.
  Relational information remains globally encoded but undergoes complete loss of
  spatiotemporal projectability, beyond the domain in which decoherence can be defin
