\subsection{Nonlocal Correlations Without Superluminality}
  \label{subsec:nonlocal-correlations-without-superluminality}

  Within the Cosmochrony framework, nonlocal quantum correlations do not arise from
  superluminal propagation of information.
  All admissible projected descriptions respect local causal constraints, and no
  measurement outcome influences another through dynamical signal exchange.
  This absence of information transfer is not enforced by a relativistic prohibition alone,
  but reflects a deeper ontological fact:
  there is no transmission of information because, at the level of the $\chi$ substrate,
  there is no fundamental separation to bridge.
  What appear as distinct outcomes correspond to local reprojections of a single underlying relational configuration.

  Correlated outcomes instead arise because spacelike separated measurements correspond
  to different local reprojections of a single non-factorizable admissible projected description.
  Such descriptions cannot be decomposed into independent subsystems without loss of global consistency.
  This failure of factorization is not accidental, but reflects the absence of an
  injective mapping between effective subsystems and underlying $\chi$ configurations.
  As a result, the factorization assumptions underlying Bell-type inequalities are
  violated, while dynamical locality and relativistic causality remain intact.
  This point is grounded in the relational formulation, where factorization is not
  fundamental but only an approximate emergent feature of certain projectable regimes
  (see Appendix~E, especially Appendix~\ref{subsec:non-factorization-and-entanglement}.

  In this perspective, quantum correlations reflect global descriptive consistency
  rather than hidden variables or pre-existing local properties.
  Measurement outcomes do not reveal predetermined values, but correspond to compatible
  local realizations selected from a shared non-factorizable descriptive structure.
  Equivalently, admissible projected descriptions determine a constrained \emph{space}
  of mutually compatible local realizations, rather than a set of pre-assigned outcomes.

  Cosmochrony therefore accounts for experimentally observed violations of Bell
  inequalities without invoking nonlocal forces, retrocausality, or hidden signal channels.
  Nonlocality appears as a structural feature of admissible projected descriptions,
  fully compatible with relativistic causal constraints.
  A compact way to state this is that the correlations are \emph{ontological}, fixed by the non-injective global
  relational structure, rather than \emph{dynamical}, mediated by any superluminal interaction.
