\subsection{Temporal Ordering and Relativistic Consistency}
  \label{subsec:temporal-ordering-and-relativistic-consistency}

  Within the Cosmochrony framework, temporal ordering is not defined by a global notion
  of simultaneity.
  It arises only at the level of effective descriptions, as an ordering of admissible
  projected events induced by constrained relaxation ordering, rather than by any
  fundamental or absolute time parameter.

  Different observers may therefore assign different temporal orderings to spacelike
  separated events within effective geometric descriptions.
  Such differences reflect the observer-dependence of spacetime slicing and have no
  impact on the underlying relational consistency of admissible projected descriptions.
  No preferred foliation, global clock, or absolute temporal structure is selected at
  the fundamental level.

  Entanglement correlations are thus fully compatible with relativistic causality.
  They do not rely on any privileged reference frame or on a globally defined temporal
  ordering.
  Instead, they arise from non-factorizable admissible projected descriptions whose
  internal relational consistency is preserved under changes of effective spacetime
  coordinates and temporal slicings.

  In this sense, relativistic covariance is maintained because the physical content of
  the theory resides in relational consistency conditions rather than in observer-dependent
  spatiotemporal labels.
  Temporal ordering remains an effective, observer-relative notion, while admissible
  correlations and their consistency relations remain invariant across all equivalent
  projected descriptions.
