\subsection{Local Slowdown of Relaxation Ordering}
  \label{subsec:local-slowdown-of-chi-relaxation}

  In the Cosmochrony framework, gravitation does not arise from a fundamental
  interaction.
  It emerges, at the level of effective descriptions, from the collective influence
  of particle-like projected configurations on admissible relaxation ordering.

  As established in Sec.~\ref{sec:particles-as-localized-excitations-of-the-chi-field},
  localized projected configurations resist admissible relaxation ordering.
  When many such configurations are present, their combined effect leads, in the
  weak-constraint regime, to an effective macroscopic reduction of the admissible
  ordering rate within projected descriptions.

  In an effective spacetime parametrization, this collective effect may be expressed
  schematically as
  \begin{equation}
    \mathcal{D}_{\mathrm{eff}} \chi_{\mathrm{eff}}
    \simeq c \left( 1 - \alpha\, \rho \right),
  \end{equation}
  where $\rho$ denotes the effective density of localized projected configurations and
  $\alpha$ encodes their average contribution to relaxation resistance.
  This expression represents a first-order approximation valid when localized
  constraints are sufficiently dilute and weakly overlapping.

  The coupling parameter $\alpha$ is not fundamental.
  It emerges as a collective property of admissible projected descriptions and depends
  on the typical structural characteristics of localized configurations.
  In the weak-field limit, dimensional consistency relates its scaling to the observed
  gravitational constant, leading to $\alpha \propto G/c^2$ when expressed in terms of
  effective inertial mass densities.
  Within this approximation, the reduction of admissible relaxation ordering admits a
  Newtonian-like interpretation in terms of an effective gravitational potential.

  Physically, this collective slowdown manifests, within effective geometric
  descriptions, as gravitational time dilation.
  No independent gravitational field or force is introduced; gravitation appears as a
  macroscopic signature of constrained admissible ordering induced by localized
  projected configurations.
