\subsection{Unified Origin of Gravitational and Electromagnetic Effects}
  \label{subsec:unified-origin-of-gravitational-and-electromagnetic-effects}

  Within the Cosmochrony framework, gravitational and electromagnetic phenomena do not
  originate from distinct fundamental entities.
  They arise as complementary effective manifestations of the same underlying
  relational substrate, once projected into regimes admitting spacetime descriptions.

  At the fundamental level, no independent gravitational or electromagnetic fields are
  postulated.
  Physical interactions appear only at the level of admissible projected descriptions,
  as different modes of constrained relational organization.

  Gravitational effects correspond to sustained and quasi-static constraints on
  admissible relaxation ordering induced by persistent localized projected
  configurations.
  When expressed in effective geometric terms, these constraints manifest as time
  dilation, attraction, and spacetime curvature.
  Gravitation therefore reflects cumulative and structural constraints on admissible
  ordering.

  Electromagnetic phenomena, by contrast, arise from time-dependent and phase-structured
  patterns within admissible projected descriptions.
  These patterns admit an effective formulation in terms of propagating, oscillatory
  descriptions carrying both attractive and repulsive interactions, consistent with the
  phenomenology of electromagnetic radiation and forces.

  In this sense, gravitation and electromagnetism differ not by their fundamental
  origin, but by the temporal organization of admissible projected descriptions.
  Gravitational phenomena correspond to quasi-static, cumulative constraints on
  ordering, whereas electromagnetic phenomena correspond to dynamically structured,
  oscillatory descriptive regimes.
  The familiar distinction between the two interactions therefore emerges at the level
  of effective descriptions, rather than from fundamentally separate fields.

  Within effective spacetime regimes, the dynamic projected patterns associated with
  electromagnetism admit a formulation equivalent to classical electrodynamics.
  An explicit derivation of the corresponding Maxwell-like equations within the
  Cosmochrony framework is provided in
  Appendix~\ref{app:emergent-electrodynamics}.
