\subsection{Collective Gravitational Coupling and Operational Geometry}
  \label{subsec:collective-gravitational-coupling-and-operational-geometry}

  The collective slowdown of $\chi$ relaxation described above not only affects the local flow of
  time but also induces an effective spatial coupling between neighboring regions of the field.
  When particle excitations are present, the resistance they impose on the relaxation of $\chi$
  modulates how efficiently variations of the field propagate across space.

  This collective behavior can be encoded in an effective local coupling between neighboring
  degrees of freedom, denoted $K_{ij}$, which characterizes the stiffness of the $\chi$ field to
  relative variations.
  In regions where $\chi$ varies smoothly, the coupling approaches a constant
  vacuum value, while strong local gradients associated with particle excitations effectively weaken
  the coupling.
  Importantly, $K_{ij}$ is assumed to be purely local and constitutive, depending only
  on neighboring values of $\chi$, rather than on any pre-existing spatial metric.

  Because no background geometry is assumed at the fundamental level, spatial distance is defined
  operationally through the propagation of $\chi$ across the network: two points are considered
  close if $\chi$-relaxation propagates efficiently between them, and distant otherwise. In the
  continuum and weak-gradient limit, this operational notion induces an effective spatial geometry,
  which can be described by a metric to leading order.

  In this sense, spacetime curvature in Cosmochrony does not arise as a primitive geometric property,
  but as a collective manifestation of how localized particle excitations modulate the propagation
  and relaxation of the $\chi$ field.

  The explicit construction of the collective coupling $K_{ij}$, the operational definition of
  distance on the discrete network, and the derivation of the effective field equations in the
  quasi-static regime are presented in Appendix~\ref{subsec:collective-coupling}.
