\clearpage
\section{Numerical Methods and Technical Supplements}
  \label{sec:appendix-technical}

  This appendix collects numerical methods, technical constructions, and auxiliary
  derivations used to explore the phenomenological consequences of the Cosmochrony
  framework.
  Its role is explicitly supportive: the material presented here does not introduce
  additional fundamental structures, nor does it modify the ontological or dynamical
  core of the theory.

  All methods described in this appendix operate within regimes where the dynamics
  of the fundamental field \(\chi\) admits an effective, discretized, or
  coarse-grained representation.
  They should therefore be understood as computational approximations and technical
  tools designed to probe the behavior of the theory, not as independent physical
  postulates.

  \paragraph{Status of the numerical constructions.}
    The fundamental formulation of Cosmochrony is relational and pre-geometric.
    It does not presuppose a background spacetime, a fixed metric, or a lattice
    structure.
    By contrast, the numerical methods employed here necessarily rely on auxiliary
    representations—such as discretized graphs, finite-difference schemes, or
    coarse-grained fields—to render the dynamics tractable.

    These representations are introduced solely for calculational convenience.
    They do not possess ontological significance and should not be interpreted as
    revealing a fundamental discreteness of the \(\chi\) field or of spacetime itself.
    Different numerical schemes may be employed without altering the conceptual
    content of the theory, provided they respect the same relaxation constraints and
    kinematic bounds.

  \paragraph{Scope of the appendix.}
    This appendix provides technical details on:
    \begin{itemize}
      \item the notion of collective gravitational coupling and the construction of an
      operational geometric description emerging from \(\chi\)-field relaxation
      (Section~\ref{subsec:collective-coupling});
      \item numerical algorithms and discretization strategies used to simulate
      \(\chi\)-field dynamics and to estimate effective model parameters;
      \item supplementary derivations and calculations that support results presented
      in the main text and in Appendices~B and~C.
    \end{itemize}

    The emphasis throughout is on internal consistency, numerical stability, and
    faithful implementation of the theoretical principles articulated in the main
    body of the work.
    No claim is made that the numerical results obtained using these methods are unique
    or exhaustive.

  \paragraph{Interpretation of numerical results.}
    Numerical simulations presented or discussed in this appendix should be regarded
    as exploratory.
    They are intended to test qualitative mechanisms—such as relaxation-driven
    emergence of geometry, solitonic stability, or scaling relations—rather than to
    produce precision predictions.

    Where numerical values are quoted, they serve as order-of-magnitude indicators or
    illustrative examples.
    Quantitative predictions suitable for direct comparison with observational data
    require more extensive simulations and systematic parameter studies, which lie
    beyond the scope of the present work.

  \paragraph{Relation to the main text.}
    The technical material collected here underpins several conceptual arguments made
    elsewhere in the paper, including:
    \begin{itemize}
      \item the emergence of effective gravitational coupling from collective
      relaxation effects;
      \item the stability and scaling properties of solitonic configurations;
      \item the qualitative cosmological and phenomenological implications discussed
      in Appendix~C.
    \end{itemize}

    Readers primarily interested in the conceptual structure of Cosmochrony may safely
    skip this appendix without loss of continuity.
    Conversely, readers interested in numerical implementation, reproducibility, or
    future computational extensions may find the detailed constructions provided here
    useful.

    \input{D-appendix-technical/D01-collective_coupling}
    \input{D-appendix-technical/D02-chi-parameters}
    \input{D-appendix-technical/D03-consistency_checks}
    \input{D-appendix-technical/D04-simulation-algorithms-for-chi-field-dynamics}
    \input{D-appendix-technical/D05-numerical-validation-of-the-chi-rightarrow-chi_eff-transition}
    \input{D-appendix-technical/D06-renormalization-hbar}
    \input{D-appendix-technical/D07-spectral_ratio_derivation}
