The stability of the $\chi$-field dynamics, governed by $\partial_t \chi = c \sqrt{1 - |\nabla \chi|^2/c^2}$,
describes the irreversible relaxation of $\chi$, where $c$ is the maximal relaxation speed.

It is essential to ensure that Cosmochrony provides a \textbf{physically consistent and predictive framework}.
Without stability, small perturbations could lead to unphysical divergences, compromising the theory's
ability to unify gravity, quantum mechanics, and cosmology.
This analysis confirms that the $\chi$ field remains well-behaved under perturbations,
validating its role as a fundamental geometric substrate for spacetime and matter.

Below, we demonstrate its stability under small perturbations, both in linear and nonlinear regimes.

\subsubsection{Linear Stability}

  Consider a spatially homogeneous base state, $\chi_0(t) = ct + \chi_{0,0}$, satisfying $\partial_t \chi_0 = c$
  . Let $\chi(x,t) = \chi_0(t) + \delta \chi(x,t)$, where $|\delta \chi| \ll |\chi_0|$
  . Substituting into the governing equation and linearizing yields:

  \[
    \partial_t \delta \chi = -\frac{|\nabla \delta \chi|^2}{2c}.
  \]

  Since the right-hand side is non-positive, \textbf{small perturbations decay over time}
  , ensuring linear stability.

\subsubsection{Nonlinear Stability}

  To assess nonlinear stability, we introduce an energy-like functional:

  \[
    E[\delta \chi] = \frac{1}{2} \int |\nabla \delta \chi|^2 \, d^3x.
  \]

  The time derivative of $E$ is:

  \[
    \frac{dE}{dt} = \int \nabla \delta \chi \cdot \nabla (\partial_t \delta \chi) \, d^3x.
  \]

  Substituting the linearized equation, we find that $E$
  is non-increasing, confirming that perturbations remain bounded. A Lyapunov functional can also be
  constructed to show that the system is \textbf{nonlinearly stable}.

\subsubsection{Special Cases}

  \begin{itemize}
    \item \textbf{Planar Waves:} For $\delta \chi = \epsilon \sin(kx - \omega t)$, the dispersion relation
    $\omega = \frac{c k^2}{2 \chi_0}$ shows that high-wavenumber perturbations are strongly damped.
    \item \textbf{Spherical Symmetry:} For $\delta \chi(r,t)$
    , the linearized equation admits diffusively decaying solutions.
  \end{itemize}

\subsubsection{Conclusion}

  The $\chi$-field dynamics are \textbf{stable}
  under both linear and nonlinear perturbations. This stability supports the viability of $\chi$
  as a fundamental field underlying spacetime, gravity, and quantum phenomena in Cosmochrony.
