\section{Radiation and Quantization}
  \label{sec:radiation-and-quantization}

  \subsection{Radiation as $\chi$–Matter Interaction}
    \label{subsec:radiation-as-$chi$matter-interaction}

  In cosmochrony, radiation does not correspond to the emission of pre-existing particles.
  Instead, it arises from the interaction between localized excitations (matter) and the surrounding $\chi$ field.

  When an excited configuration interacts with $\chi$, part of the excitation may detach as a propagating crest.
  This process is stochastic, reflecting local fluctuations of $\chi$, and gives rise to radiation.

\subsection{Emergence of Photons}
    \label{subsec:emergence-of-photons}

  Photons are not fundamental entities in this framework.
  They correspond to transient, propagating disturbances of $\chi$ generated during interactions with matter.

  Prior to detection or emission, no localized photon exists.
  Quantization appears only at the moment of interaction, when continuous $\chi$
  dynamics produces discrete energy transfer.

  Although propagating electromagnetic waves correspond to continuous $\chi$
  -disturbances, localized photon-like excitations only emerge during interactions with matter. In a double-slit
  experiment, the interference pattern arises from the continuous wave nature of $\chi$
  , while individual detection events correspond to interaction-induced localizations. This duality explains why
  photons exhibit both wave-like and particle-like behavior depending on the measurement context, without
  invoking wavefunction collapse as a fundamental process.

  \subsection{Geometric Origin of $E = h\nu$}
    \label{subsec:energy-frequency-radiation}

    This section develops, in the context of radiation processes, the energy--frequency relation
    introduced earlier in Section~\ref{subsec:energy-frequency-solitons} for localized excitations
    of the $\chi$ field.

    In Cosmochrony, radiative events correspond to the partial release of relaxation potential stored
    in localized matter excitations into propagating disturbances of the $\chi$ field.
    The energy transferred during such an event is associated with the local curvature and oscillatory
    structure of the emitted $\chi$ disturbance.
    Higher-frequency disturbances correspond to tighter curvature and therefore to a larger amount of
    relaxation potential being redistributed.

    The Planck relation
    \begin{equation}
      E = h \nu
    \end{equation}
    is thus interpreted as an effective geometric proportionality between the frequency of a propagating
    $\chi$ disturbance and the amount of relaxation potential released during an interaction.
    In this framework, the constant $h$ does not represent a fundamental quantum postulate but an effective
    conversion factor relating oscillation frequency to curvature-based energy within the $\chi$ field.

    Importantly, this interpretation does not constitute a derivation of $h$ from first principles.
    Rather, it provides a geometric reading of why energy transfer in radiative processes scales linearly
    with frequency across a wide range of phenomena.
    In the photoelectric effect, the threshold frequency $\nu_0$ corresponds to the minimal curvature
    required to liberate an electron soliton from its binding configuration, while the linear dependence
    on $\nu$ reflects the amount of relaxation potential carried by the emitted $\chi$ disturbance.

    This perspective preserves the empirical success of quantum electrodynamics while situating
    quantization as an emergent feature of interaction dynamics within a continuously evolving field,
    rather than as a fundamental discreteness imposed at the outset.

\subsection{Vacuum Fluctuations and the Casimir Effect}
    \label{subsec:vacuum-fluctuations-and-the-casimir-effect}

  Vacuum fluctuations correspond to stochastic variations of $\chi$ in the absence of localized excitations.
  Boundary conditions imposed by matter constrain these fluctuations, altering the local spectrum of allowed modes.

  The Casimir effect arises naturally as a pressure difference resulting from modified $\chi$
  dynamics between closely spaced boundaries.

\subsection{Weakly Interacting Radiation}\label{subsec:weakly-interacting-radiation}

  Disturbances with minimal curvature, such as low-frequency electromagnetic waves or neutrino-like excitations,
  interact weakly with matter.
  Their near-planar structure reduces the probability of producing localized energy transfer.

  This explains the transparency of the vacuum to radiation and the weak interaction cross sections of certain
  particles.

\subsection{Summary}
  \label{subsec:summary3}

  Radiation and quantization arise from the interaction between matter excitations and the $\chi$ field.
  Photons emerge during interactions rather than existing as independent entities, and quantization reflects
  geometric constraints of $\chi$ dynamics.
