\subsection{Ontological Parsimony and the Metric}
  \label{subsec:ontological-parsimony-and-the-metric}

  As emphasized throughout the preceding discussion, the spectral operator relevant
  for mass generation is defined independently of any emergent geometric or
  dynamical description.
  A potential criticism of Cosmochrony is that it merely replaces one geometric
  structure (the spacetime metric) with another fundamental entity (the $\chi$
  field).
  This subsection clarifies why this replacement constitutes genuine ontological
  simplification rather than a relabeling of degrees of freedom.

  \paragraph{Distinction from metric-based theories.}
    In General Relativity and related metric frameworks:
    \begin{itemize}
      \item the metric $g_{\mu\nu}$ is a fundamental tensor field with ten independent
      components,
      \item spacetime curvature is treated as a primitive geometric property,
      \item matter and energy are conceptually distinct from geometry and are coupled
      to it via the stress--energy tensor.
    \end{itemize}

    In Cosmochrony:
    \begin{itemize}
      \item only a single pre-geometric relational substrate $\chi$ is taken as
      fundamental,
      \item the spacetime metric arises as an effective, coarse-grained descriptor of
      $\chi$ relaxation dynamics and carries no independent ontological status,
      \item matter, energy, and geometry correspond to distinct regimes and invariant
      patterns of the same underlying relational structure.
    \end{itemize}

  \paragraph{Operational distinguishability.}
    The two frameworks are operationally distinct rather than notationally
    equivalent:
    \begin{enumerate}
      \item \textbf{Degrees of freedom:}
      General Relativity propagates two tensorial gravitational-wave polarizations
      derived from the metric structure.
      In Cosmochrony, the fundamental dynamics involves no independent geometric
      degrees of freedom; only relational variations of $\chi$ evolve, with effective
      tensorial behavior emerging solely at the macroscopic, coarse-grained level.

      \item \textbf{Singularities:}
      In metric theories, singularities correspond to divergences of the fundamental
      geometric structure.
      In Cosmochrony, apparent singular behavior signals the breakdown of the
      effective geometric description, while the underlying $\chi$ substrate remains
      well-defined.

      \item \textbf{Quantum regime:}
      Quantizing General Relativity requires the quantization of the metric itself
      (e.g., via the Wheeler--DeWitt equation).
      In Cosmochrony, quantization applies only to effective excitations and
      fluctuations of $\chi$ within regimes where a spacetime description is already
      valid; the $\chi$ substrate itself is not treated as a conventional quantum
      field propagating on spacetime~\cite{Rovelli1997}.
    \end{enumerate}

  \paragraph{Ontological economy.}
    From the perspective of ontological parsimony, Cosmochrony achieves unification
    through reduction rather than proliferation:
    \begin{align}
      \text{Standard approach:} \quad
      & g_{\mu\nu} \; (\text{geometry}) + \psi \; (\text{matter}) + \Lambda \;
      (\text{dark energy}), \\
      \text{Cosmochrony:} \quad
      & \chi \; (\text{single relational substrate})
      \;\longrightarrow\; \{\text{spacetime}, \text{matter}, \text{expansion}\}.
    \end{align}

    This reduction does not merely rephrase existing structures, but provides an
    explanatory compression in which multiple physical notions arise as effective
    manifestations of a single underlying dynamical principle.
