\subsection{Ontological Parsimony and the Metric}
    \label{subsec:ontological-parsimony-and-the-metric}

    A potential criticism of Cosmochrony is that it merely replaces one geometric structure (the metric) with another
    (the $\chi$ field).
  This section addresses why this replacement constitutes genuine ontological progress rather than relabeling.

  \paragraph{Distinction from metric theories.}
    In General Relativity and its extensions:
    \begin{itemize}
      \item The metric $g_{\mu\nu}$ is a fundamental tensor field with 10 independent components.
      \item Spacetime curvature is a primitive geometric property.
      \item Matter and energy are conceptually distinct from geometry, coupled via the stress-energy tensor.
    \end{itemize}

    In Cosmochrony:
    \begin{itemize}
      \item Only the scalar field $\chi$ (1 component) is fundamental.
      \item The metric is a derived effective description, not an independent dynamical entity.
      \item Matter, energy, and geometry are unified as different manifestations of $\chi$ configurations.
    \end{itemize}

  \paragraph{Operational distinguishability.}
    The frameworks are operationally distinct:
    \begin{enumerate}
      \item \textbf{Degrees of freedom:} GR propagates 2 gravitational wave polarizations from 10 metric components.
      Cosmochrony propagates perturbations of 1 scalar field, with effective tensorial structure emerging only macroscopically.

      \item \textbf{Singularities:} GR singularities (where $g_{\mu\nu}$ diverges) are ontological.
      In Cosmochrony, apparent singularities mark the breakdown of the effective metric description, while $\chi$ remains well-defined.

      \item \textbf{Quantum regime:} Quantizing GR requires quantizing the metric (Wheeler-DeWitt equation).
      Quantizing Cosmochrony requires only quantizing $\chi$, with spacetime emerging from quantum $\chi$ configurations.
    \end{enumerate}

  \paragraph{The Occam's razor argument.}
    Cosmochrony achieves unification through reduction:
    \begin{align}
      \text{Traditional:} \quad & g_{\mu\nu} \,(\text{10 DOF}) + \psi \,(\text{matter}) + \Lambda \,(\text{dark energy}) \\
      \text{Cosmochrony:} \quad & \chi \,(\text{1 DOF}) \longrightarrow \{\text{spacetime}, \text{matter}, \text{expansion}\}
    \end{align}

    This represents genuine explanatory compression, not mere reformulation.
