\subsection{Status of the Wavefunction}
  \label{subsec:status-of-the-wavefunction}

  In standard quantum mechanics, the wavefunction $\psi$ is a complex-valued object
  defined on configuration space, whose ontological status remains debated.
  Operationally, $|\psi|^2$ encodes measurement probabilities via the Born rule, while
  $\psi$ itself does not correspond to a physical field propagating in spacetime.

  Within the Cosmochrony framework, the quantum wavefunction is not identified with any
  fundamental physical entity.
  It arises at the level of effective descriptions as a statistical encoding of the
  set of admissible local reprojections compatible with a given non-factorizable
  projected configuration.

  The wavefunction therefore does not represent an underlying structure or hidden
  dynamics.
  It summarizes, in compact mathematical form, the relative accessibility of different
  local outcomes once a global descriptive coherence has been established.
  Its complex phase encodes relational constraints between alternative local
  descriptions, while its modulus determines the statistical weight of accessible
  reprojections.

  As an illustrative example, the hydrogen atom wavefunctions
  $\psi_{nlm}(r,\theta,\phi)$ do not correspond to localized structures or solitonic
  configurations at a fundamental level.
  They represent stationary admissible projected descriptions characterized by
  specific symmetry and stability properties.
  The probability density $|\psi|^2$ reflects the relative frequency with which
  localized reprojections occur in effective geometric descriptions, while energy
  quantization arises from the discrete admissibility conditions imposed on stationary
  descriptive regimes.

  In this perspective, the wavefunction is neither a physical field nor a direct image
  of an underlying substrate.
  It is a derived statistical object, encoding the structure of admissible projected
  descriptions and the probabilities of their local realization within spacetime.
