\subsection{Scope and Limitations}
  \label{subsec:scope-and-limitations}

  Cosmochrony does not aim to replace quantum mechanics as a predictive or
  computational framework.
  All standard quantum-mechanical formalisms, including operator methods, path
  integrals, and perturbative techniques, remain valid and unchanged within their
  established domains of applicability.

  The contribution of Cosmochrony is interpretative and unificatory.
  It provides a coherent pre-geometric and relational account of the structural
  origin of quantum phenomena, clarifying the ontological status of the wavefunction,
  quantization, measurement, and nonlocal correlations, without altering any
  experimentally verified predictions.

  Within this framework, quantum mechanics is understood as an effective theory
  governing admissible projected descriptions in regimes where linearity,
  localization, and approximate factorization hold.
  Cosmochrony does not introduce new degrees of freedom, hidden variables, or
  modifications of quantum dynamics.

  A complete and systematic formal correspondence between the relational $\chi$
  substrate and the operator-based structures of quantum theory---including a systematic derivation
  of Hilbert space, observables, and evolution operators---lies beyond the scope of
  the present work and is left for future investigation.

  Accordingly, the present framework should be regarded as a foundational
  reconstruction rather than a competing physical theory, intended to clarify the
  conceptual origin, internal coherence, and domain of validity of
  quantum-mechanical descriptions.
