\subsection{Spectral Stability and the Emergence of \(\hbar_{\mathrm{eff}}\)}
\label{sec:hbar_eff_derivation}

In Cosmochrony, the effective Planck constant \(\hbar_{\mathrm{eff}}\) is not
introduced as a fundamental quantum postulate.
Instead, it emerges as a scaling parameter linking spectral stability of
\(\chi\)-field solitons to effective spacetime observables.

\subsubsection*{Fundamental scales of the \(\chi\) dynamics}

  The \(\chi\) field is characterized by three independent dynamical scales:
  \begin{itemize}
    \item \(K_0\): maximal relaxation stiffness, with dimensions \([L^{-2}]\),
    \item \(\chi_c\): correlation length at which solitonic configurations stabilize,
    \item \(c\): maximal relaxation speed.
  \end{itemize}

  From these, one may define a natural unit of action associated with the relaxation
  dynamics,
  \begin{equation}
    \hbar_{\chi}
    \;\equiv\;
    \frac{c^3}{K_0 \chi_c} ,
  \end{equation}
  which has the dimensions of action and is independent of the standard Planck
  constant.

\subsubsection*{Spectral origin of effective quantization}

  Quantization in Cosmochrony follows from the discrete spectrum of the stability
  operator \(\Delta_G^{(0)}\).
  For a solitonic excitation with eigenvalue \(\lambda_n\), the characteristic
  frequency of small oscillations scales as
  \begin{equation}
    \nu_n \;\sim\; \frac{c}{\chi_c}\,\sqrt{\lambda_n}\,\mathcal{N}_n^{1/2} .
  \end{equation}

  Identifying the rest energy with the product of this frequency and an effective
  action scale yields
  \begin{equation}
    E_n = \hbar_{\mathrm{eff}}\,\nu_n ,
  \end{equation}
  from which \(\hbar_{\mathrm{eff}}\) emerges as a geometric and spectral quantity,
  not as an independent constant.

\subsubsection*{Regime-dependent scaling}

  The effective value of \(\hbar_{\mathrm{eff}}\) depends on the scale at which the
  system is probed.
  In regimes where the characteristic spacetime scale
  \(\ell_{\mathrm{spacetime}}\) is comparable to \(\chi_c\),
  \begin{equation}
    \hbar_{\mathrm{eff}} \approx \hbar_{\chi} ,
  \end{equation}
  recovering standard quantum behavior.

  At macroscopic scales \(\ell_{\mathrm{spacetime}} \gg \chi_c\),
  \begin{equation}
    \hbar_{\mathrm{eff}}
    \approx
    \hbar_{\chi}
    \left( \frac{\chi_c}{\ell_{\mathrm{spacetime}}} \right)^2 ,
  \end{equation}
  leading to a strong suppression of quantum effects and the emergence of classical
  behavior.

\subsubsection*{Consistency with quantum phenomenology}

  In the microscopic regime, where \(\hbar_{\mathrm{eff}} \approx \hbar\), standard
  quantization relations
  \(E = \hbar \nu\) are recovered as effective descriptions.
  This agreement is not postulated but follows from the scaling behavior of
  \(\hbar_{\mathrm{eff}}\) once the projected regime matches laboratory scales.

  \paragraph{Numerical constraints.}

    Reproducing particle-scale quantum behavior requires
    \begin{equation}
      K_0 \chi_c^2 \sim \hbar ,
    \end{equation}
    which constrains the admissible values of the relaxation stiffness and correlation
    length.
    These constraints are consistent with soliton stability and do not require fine
    tuning.

  \paragraph{Summary.}

    Within Cosmochrony, both inertial mass and effective quantization emerge from the
    same spectral stability structure of the \(\chi\) relaxation dynamics.
    The Planck constant appears not as a fundamental input, but as a scale-dependent
    parameter encoding the projection from relational dynamics to spacetime-based
    observables.
