\section{Cosmological and Observational Implications of Cosmochrony}
  \label{sec:appendix-cosmo}

  This appendix examines the cosmological and observational implications of the
  Cosmochrony framework.
  Its purpose is not to construct a fully parameterized cosmological model, nor to
  perform precision fits to existing datasets, but to establish conceptual and
  phenomenological consistency with key observations and to identify distinctive
  signatures that may allow the framework to be tested or falsified.

  All results presented here should be understood as consequences of the same
  underlying scalar relaxation dynamics governing the \(\chi\) field.
  No additional cosmological degrees of freedom are introduced beyond those already
  discussed in the main body of the work.
  Standard cosmological observables arise as effective descriptions once a smooth
  geometric projection of \(\chi\) becomes applicable.

  The appendix is organized as follows:
  \begin{itemize}
    \item Section~\ref{app:lowell_attenuation} analyzes the spectrum of large-scale
    fluctuations of the projected field \(\chi_{\mathrm{eff}}\) and their imprint on
    the cosmic microwave background, with particular emphasis on the suppression of
    low-\(\ell\) multipoles.
    \item Section~\ref{subsec:cosmochrony_horizon_flatness} shows how the horizon and
    flatness problems are resolved dynamically in Cosmochrony, without invoking an
    inflationary phase or fine-tuned initial conditions.
    \item Section~\ref{app:hubble_tension} discusses the evolution of the effective
    Hubble parameter \(H(z)\) and its implications for the observed Hubble tension,
    highlighting how deviations from standard expansion histories may arise from
    relaxation dynamics.
    \item Section~\ref{subsec:observational-estimates} provides order-of-magnitude
    estimates of the characteristic \(\chi\)-field parameters and relates them to
    observable cosmological quantities, without assuming a specific cosmological
    parameter set.
    \item Section~\ref{subsec:phenomenology} explores broader phenomenological
    consequences, including modified gravitational-wave propagation, MOND-like
    effective dynamics at galactic scales, and lensing anomalies.
  \end{itemize}

  Throughout this appendix, the emphasis is placed on identifying robust qualitative
  and semi-quantitative features that follow generically from the Cosmochrony
  framework.
  Where numerical estimates are provided, they are intended as consistency checks
  and scaling guides rather than as precision predictions.

  Taken together, these results demonstrate that Cosmochrony is compatible with the
  broad structure of contemporary cosmological observations while predicting
  systematic deviations from standard \(\Lambda\)CDM expectations.
  These deviations define concrete targets for future observational tests and
  numerical investigations.

  \input{C-appendix-cosmo/C01-chi_cmb_spectrum}
  \input{C-appendix-cosmo/C02-cosmochrony_horizon_flatness}
  \input{C-appendix-cosmo/C03-hubble_tension}
  \input{C-appendix-cosmo/C04-relation-to-observational-units-and-numerical-estimates}
  \input{C-appendix-cosmo/C05-phenomenological-implications}
  \input{C-appendix-cosmo/C06-spectral-gravitational-toy-model}
