\subsection{Resolution of the Horizon and Flatness Problems Without Inflation}
  \label{subsec:cosmochrony_horizon_flatness}

  In standard cosmology, the horizon and flatness problems arise from extrapolating
  a spacetime-based notion of causality and geometry back to the earliest stages of
  cosmic evolution.
  Within this framework, regions of the universe that appear widely separated today
  should not have been in causal contact, and the near-flatness of spatial geometry
  requires fine-tuned initial conditions.
  Inflation addresses these issues by postulating a brief phase of accelerated
  expansion in a pre-existing metric background.

  Cosmochrony adopts a fundamentally different standpoint.
  Spacetime geometry, causal structure, and metric notions of distance are not
  assumed to be fundamental.
  They emerge only at a later stage, as effective descriptions of the relaxation
  dynamics of the scalar field \(\chi\).
  As a result, the assumptions underlying the horizon and flatness problems do not
  apply at the fundamental level.

  \paragraph{Horizon problem: pre-geometric connectivity.}
    In Cosmochrony, large-scale correlations do not need to be established through
    signal propagation within spacetime.
    Instead, they originate from the fact that \(\chi\) constitutes a single,
    globally connected dynamical substrate whose relaxation precedes the emergence of
    any effective spacetime description.

    At early stages, before a metric notion of causality becomes meaningful, the
    configuration of \(\chi\) is defined globally.
    Regions that later appear causally disconnected in the emergent spacetime may
    therefore share correlated configurations inherited from earlier phases of the
    relaxation process.

    In this sense, Cosmochrony replaces inflationary causal contact with
    \emph{pre-geometric connectivity}:
    correlations are established at the level of the fundamental field itself, rather
    than through superluminal expansion or specially prepared initial conditions on a
    metric background.

  \paragraph{Flatness problem: relaxation toward geometric uniformity.}
    The flatness problem is addressed through the same underlying mechanism.
    In Cosmochrony, effective spatial curvature reflects large-scale gradients and
    inhomogeneities in the relaxation rate of the projected field
    \(\chi_{\mathrm{eff}}\).
    As relaxation proceeds, configurations with large curvature gradients are
    dynamically disfavored, since they correspond to sustained resistance to global
    relaxation.

    As a consequence, near-flat spatial geometry emerges as a natural attractor of the
    relaxation dynamics.
    Curvature dilution does not require exponential expansion or fine-tuning of
    initial curvature parameters.
    It reflects the tendency of the \(\chi\) field to minimize large-scale geometric
    tension as it approaches a homogeneous relaxation state.

    This mechanism operates independently of any inflationary phase and does not rely
    on a specific initial curvature value.

  \paragraph{Implications for primordial correlations.}
    Because large-scale coherence arises from the global organization of \(\chi\)
    rather than from the amplification of quantum vacuum fluctuations, Cosmochrony
    does not predict exact scale invariance at the largest wavelengths.
    Instead, the longest-wavelength modes are subject to global relaxation
    constraints, which may lead to deviations from scale invariance at the lowest
    multipoles of the cosmic microwave background.

    Such deviations are interpreted as structural tendencies rather than sharp
    predictions.
    They provide a qualitative distinction from inflation-based scenarios and
    motivate the phenomenological analysis of low-\(\ell\) CMB anomalies discussed in
    Section~\ref{app:lowell_attenuation}.

  \paragraph{Status and limitations.}
    The arguments presented here establish that the horizon and flatness problems do
    not arise as fundamental inconsistencies within the Cosmochrony framework.
    They are artifacts of applying metric-based reasoning beyond its domain of
    validity.

    A quantitative derivation of primordial correlation functions and power spectra,
    including detailed predictions for CMB anisotropies, requires dedicated numerical
    simulations of the \(\chi\)-field relaxation dynamics and lies beyond the scope of
    the present work.

    Nevertheless, Cosmochrony provides a conceptually coherent, inflation-free
    resolution of large-scale causal coherence and near-flat spatial geometry, rooted
    in the pre-geometric dynamics of a single scalar field.
